Surrogate model for an aligned-spin effective one body waveform model of binary neutron star inspirals using Gaussian process regression

### 2018-12-02 paper/paper_no_event.tex
Paper: changes in II C
- renamed to 'waveform termination'
- effective post-merger rather than post-merger model
- changed the 20M entering the definition of tfreq to 12M

\subsection{Post-merger waveform}

We now look for a suitable time $t_{\rm match}$ to stop the numerical integration and match with a post-merger waveform. In the point-mass model, \texttt{SEOBNRv4}, nonquasicircular corrections to the inspiral-plunge signal guarantee that the waveform peaks, and $t_{\rm match}$ is chosen to be the time of the peak amplitude $t_{\rm peak}^{\rm amp}$. Here, we do not use nonquasicircular corrections tuned to BNS simulations, so we have less control over the behavior of Eq.~(\ref{hlminspplunge}) in the late inspiral. We therefore choose the following definition for $t_{\rm match}$. Let $t_{\rm peak}^{\rm amp}$ be the earliest time when the amplitude $|h_{22}^{\textrm{insp-plunge}}|$ of the (2, 2) mode peaks, and let $t_{\rm peak}^{\rm freq}$ be the time when the frequency $\omega_{22}^{\rm insp-plunge}$ of the (2, 2) mode peaks. Then,
\begin{equation}
t_{\textrm{match}} = \min \left(t_{\textrm{peak}}^{\textrm{amp}},t_{\textrm{peak}}^{\textrm{freq}}\right)\,.\label{tmatch}
\end{equation}
For some waveforms, the amplitude does not peak. In this case, we choose $t_{\rm peak}^{\rm amp}$ to be the earliest time when the slope of the amplitude $\partial_t |h_{22}^{\rm insp-plunge}|$ reaches a minimum after having reached a peak.

@@

In BNS simulations, the waveform rapidly tapers to zero amplitude in 1--2 gravitational-wave cycles after reaching peak amplitude. The BNS system then either undergoes prompt collapse or post-merger oscillations of the remnant. We model this tapering in terms of analytic functions for the amplitude $A^{\textrm{post-mrg}}(t)$ and phase $\phi^{\textrm{post-mrg}}(t)$ such that $h_{22}^{\textrm{post-mrg}}(t) = A^{\textrm{post-mrg}}(t) \exp{[i\phi^{\textrm{post-mrg}}(t)]}$.

@@

For the phase, we smoothly extend the waveform frequency such that it agrees with the inspiral frequency at a time $t_{\rm freq} = t_{\rm match} - 20M$ before the matching time $t_{\rm match}$, but then stretches out the frequency evolution such that it only approaches $\omega_{\rm match}$ asymptotically. We define this frequency evolution as
\begin{align}
\omega(t) = \omega_{\rm match} - \Delta\omega \exp\left[ -(t - t_{\rm freq}) / (20M) \right],
\end{align}
where $\Delta\omega = \omega_{\rm match} - \omega_{\rm freq}$. Integrating the frequency, and requiring continuity at $t_{\rm freq}$, results in the final expression for the phase
\begin{align}
\phi_{22}(t) = \left\{\begin{array}{ll}
\phi_{22}^{\textrm{insp-plunge}}(t), & t \leq t_{\rm freq} \,, \\
\phi_{\rm freq} + \omega_{\rm match}(t - t_{\rm freq}) + 20M \Delta\omega \left\{ \right. & \\
  \left. \exp\left[ -(t - t_{\rm freq}) / (20M) \right] - 1\right\} \,,& t > t_{\rm freq} \, .
\end{array}\right.
\end{align}

Although this post-merger model has not been fit to NR BNS simulations, it is in reasonable qualitative agreement with the post-merger behavior of the two equal-mass nonspinning NR BNS simulations that were analyzed in Ref.~\cite{Hinderer:2016eia}. A more sophisticated, NR-informed model of the post-merger emission will be part of future investigations.

>>>>

\subsection{Waveform termination}

We now look for a suitable time $t_{\rm match}$ to stop the numerical integration and match with an effective post-merger waveform. In the point-mass model, \texttt{SEOBNRv4}, nonquasicircular corrections to the inspiral-plunge signal guarantee that the waveform peaks, and $t_{\rm match}$ is chosen to be the time of the peak amplitude $t_{\rm peak}^{\rm amp}$. Here, although we include nonquasicircular corrections, they are computed as in the BBH case and are not tuned to BNS simulations, so we have less control over the behavior of Eq.~(\ref{hlminspplunge}) in the late inspiral. We therefore choose the following definition for $t_{\rm match}$. Let $t_{\rm peak}^{\rm amp}$ be the earliest time when the amplitude $|h_{22}^{\textrm{insp-plunge}}|$ of the (2, 2) mode peaks, and let $t_{\rm peak}^{\rm freq}$ be the time when the frequency $\omega_{22}^{\rm insp-plunge}$ of the (2, 2) mode peaks. Then,
\begin{equation}
t_{\textrm{match}} = \min \left(t_{\textrm{peak}}^{\textrm{amp}},t_{\textrm{peak}}^{\textrm{freq}}\right)\,.\label{tmatch}
\end{equation}
For some waveforms, the amplitude does not peak. In this case, we choose $t_{\rm peak}^{\rm amp}$ to be the earliest time when the slope of the amplitude $\partial_t |h_{22}^{\rm insp-plunge}|$ reaches a minimum after having reached a peak.

@@

In BNS simulations, the waveform rapidly decreases in amplitude in 1--2 gravitational-wave cycles after reaching peak amplitude. The BNS system then either undergoes prompt collapse or post-merger oscillations of the remnant. We do not attempt to model post-merger oscillations, and instead model an approximate peak emission and a subsequenct tapering to zero. This signal will be represented in terms of analytic functions for the amplitude $A^{\textrm{post-mrg}}(t)$ and phase $\phi^{\textrm{post-mrg}}(t)$ such that $h_{22}^{\textrm{post-mrg}}(t) = A^{\textrm{post-mrg}}(t) \exp{[i\phi^{\textrm{post-mrg}}(t)]}$.

@@

For the phase, we smoothly extend the waveform frequency such that it agrees with the inspiral frequency at a time $t_{\rm freq} = t_{\rm match} - 12M$ before the matching time $t_{\rm match}$, but then stretches out the frequency evolution such that it only approaches $\omega_{\rm match}$ asymptotically. We define this frequency evolution as
\begin{align}
\omega(t) = \omega_{\rm match} - \Delta\omega \exp\left[ -(t - t_{\rm freq}) / (12M) \right],
\end{align}
where $\Delta\omega = \omega_{\rm match} - \omega_{\rm freq}$. Integrating the frequency, and requiring continuity at $t_{\rm freq}$, results in the final expression for the phase
\begin{align}
\phi_{22}(t) = \left\{\begin{array}{ll}
\phi_{22}^{\textrm{insp-plunge}}(t), & t \leq t_{\rm freq} \,, \\
\phi_{\rm freq} + \omega_{\rm match}(t - t_{\rm freq}) + 12M \Delta\omega \left\{ \right. & \\
  \left. \exp\left[ -(t - t_{\rm freq}) / (12M) \right] - 1\right\} \,,& t > t_{\rm freq} \, .
\end{array}\right.
\end{align}

Although this effective post-merger model has not been fit to NR BNS simulations, it is in reasonable qualitative agreement with the post-merger behavior of the two equal-mass nonspinning NR BNS simulations that were analyzed in Ref.~\cite{Hinderer:2016eia}. A more sophisticated, NR-informed model of the post-merger emission will be part of future investigations.
